\SW: New Lens Candidates from the CFHT Legacy Survey

### 2014-06-04 doc/sw-cfhtls.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Space Warps System Paper
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\title[\sw]
{\SW: Crowd-sourcing the Discovery of Gravitational Lenses}

\author[Marshall et al.]{%
  \input{sw-system-authors.tex}
}

@@

\sw is a web-based service that enables the discovery of strong gravitational
lenses in wide-field imaging surveys by large numbers of people. Carefully
produced color composite images are displayed to volunteers via a
classification interface which records their estimates of the positions of
candidate lensed features. Simulated lenses, and expert-classified non-lenses,
are inserted into the image stream at random intervals; this training set is
used to give the volunteers feedback on their performance, and to estimate a
dynamically-updated probability for any given image to contain a lens. Low
probability systems are retired from the site periodically, concentrating the
sample towards a set of candidates; this ``stage 1'' set is then re-classified
by the volunteers in a second  refinement stage. Analyzing the classification
of the training set, we predict that the first stage alone should yield a
sample that is C\% complete, while leading to the rejection of R\% of the
initial target sample. Having divided the 150 square degree CFHTLS imaging
survey into 430000 overlapping 70 by 70 arcminute tiles and displayed them on
the site, we were joined by 33000 volunteers who contributed X million image
classifications over the course of N months. The sample was reduced to 3500
stage 1 candidates; these were then refined to yield a sample of 1400
candidates rankable by their stage 2 probability. We expect this sample to be
X\% complete and Y\% pure at a threshold of 95\% classification probability.
We find that, on average, and given the assumptions we make in our analysis,
we need 9 classifications per image during the first stage, X in the second.
We estimate the mean information contributed per person to be X bits, over a
session lasting, on average, N classifications per volunteer, and present the
highly skewed distributions of these quantities. We comment on the scalability
of the \sw system to the wide field survey era, and its potential to operate
beyond its design as a supervised classification system.

@@

% Scientific motivation. Applications of lenses: group-scale arcs,
% galaxy-galaxy lenses, lensed quasars.

Strong gravitational lensing -- the formation of multiple, magnified images of
background objects due to the deflection of light by  massive foreground
objects -- is a very powerful astrophysical tool, enabling a wide range of
science projects. The image separations and distortions provide information
about the mass distribution in the lens \citep[e.g.][]{AugerEtal2010,
SonnenfeldEtal2012,SonnenfeldEtal2013}, including on sub-galactic scales
\citep[e.g.][]{Dalal+Kochanek2002,VegettiEtal2010,HezavehEtal2013}. Any
strong lens can provide magnification of a factor of 10 or more, providing a
deeper, higher resolution view of the distant universe through these ``cosmic
telescopes'' \citep[e.g.][]{StarkEtal2008,NewtonEtal2011}. Lensed quasars enable
cosmography via the time delays between the multiple images' lightcurves
\citep[e.g.][]{TewesEtal2013,SuyuEtal2013}, and study of the accretion disk itself
through the microlensing effect \citep[e.g.][]{PoindexterEtal2008}. All of these
science projects would benefit from being able to draw from a larger sample of
lenses.

In the last decade the numbers of detections of these rare cosmic alignments
has increased by an order of magnitude, thanks to wide field surveys such as
CLASS \citep[e.g.]{BrowneEtal2003}, SDSS \citep[e.g.][]
{BoltonEtal2006,AugerEtal2010b,TreuEtal2011,InadaEtal2012}, CFHTLS
\citep[e.g.][]{MoreEtal2012,GavazziEtal2014}, Herschel
\citep[][]{NegrelloEtal2014} and SPT \citep[e.g.][]{VieiraEtal2013}, among
others.  As the number of known lenses has increased, new types have been
discovered, leading to entirely new investigations. Compound lenses
\citep{GavazziEtal2008,CollettEtal2012} and lensed supernovae
\citep{QuimbyEtal2014} are good examples of this.

% Problem of rarity. Imaging surveys. Problem of purity/false positives.
% Review of progress to date. Methods in SL2S, SQLS. Contrast with SLACS.

Because they are rare, strong lenses are expensive to find. The most efficient
searches to date have made use of relatively clean signals such as the
presence of emission or absorption features at two distinct redshifts in the
same optical spectrum \citep[e.g.][]{BoltonEtal2004}, or the strong
``magnification bias'' towards detecting strongly-lensed sources in the sub-mm
waveband \citep[e.g.][]{NegrelloEtal2010}. Such searches have to be efficient,
because they require expensive high resolution imaging follow-up; consequently
they have so far produced yields in the tens to hundreds. An alternative approach
is to search images of sufficiently high resolution and color contrast, and
confirm the systems as gravitational lenses by modeling the survey data
themselves \citep[][]{MarshallEtal2009}. Several square degrees of HST images
have been searched, yielding several tens of galaxy-scale lenses
\citep[e.g.][]{MoustakasEtal2007,FaureEtal2008,Jackson2008,MoreEtal2012,
PawaseEtal2014}.
Similarly, searches of over a hundred square degrees of CFHT Legacy Survey
ground-based imaging, also with sub-arcsecond image quality, have revealed a
smaller number of wider image separation group-scale systems
\citep[e.g.][]{CabanacEtal2007,MoreEtal2012}. Detecting galaxy-scale lenses from
the ground is hard, but feasible albeit lower efficiency and requiring HST or
spectroscopic follow-up to confirm the candidates as lenses
\citep[e.g.][]{GavazziEtal2014}.

% Scaling to wide field era. Automated methods: problems. Need for good
% training sets. Need for quality control: always present.

How can we scale these lens searches up to imaging surveys covering a hundred
times the sky area, such as the almost-all sky surveys planned with LSST and
Euclid, while reducing our dependence on expensive follow-up confirmation
observations? There are two approaches to detecting lenses in imaging surveys.
The first one is robotic: automated analysis of object catalogs and/or the
survey images. The candidate samples produced by these methods have, to date,
not been of high purity  \citep[see
e.g.][]{MarshallEtal2009,MoreEtal2012,GavazziEtal2014}, with visual inspection by
teams of humans still required to narrow down the robotically-generated
samples. In this approach, the image data may or may not be explicitly
modelled by the robots as if it contained a gravitational lens, but the visual
inspection can be thought of as a ``mental modeling'' step. Systems classified
by an inspector to be good lens candidates are deemed as such because the
features in the image can be explained by a model of what gravitational lenses
do contained in the inspector's brain. The second approach simply cuts out the
robot middleman: \citet{FaureEtal2008,Jackson2008} and \citet{PawaseEtal2014}
all performed entirely visual searches for lenses in HST imaging.

Visual image inspection seems, at present, unavoidable at some level when
seraching for gravitational lenses. The technique has some drawbacks, however.
First is that humans are only humans, and they make mistakes. The solution to
this is to operate in teams, providing multiple classifications of the same
images in order to catch errors and correct them. Second, and relatedly, is
that humans get tired. With a well-designed classification interface, a human
might be able to inspect images at a rate of one astronomical object per
second (provided the majority are indeed uninteresting). At $10^4$ massive
galaxies, and 10 lenses, per square degree, visual lens searches in good
quality imaging data are limited to a few square degrees per inspector per
day. Scaling to thousands of square degrees therefore means either robotically
reducing the number of targets for inspection, or increasing the number of
inspectors, or both.

For example, a $10^4$ square degree survey containing $10^8$
photometrically-selected massive galaxies and $10^5$ lenses could only be
searched by 10 inspectors at a mean rate of 1 galaxy per second and 10
inspections per galaxy in about 14 years. Reducing the inspection time by a
factor of 400 to two weeks would require a robot to reduce the target sample
to 25 per square degree. However, at this point the required purity, 40\%,
would very likely require the average classification time per object to be
more like 10 seconds per object. Hiring 10 inspectors to assess complex images
full time full time for five months may not be the most cost-effective or
reliable strategy. Alternatively, a team of $10^6$ inspectors could, in
principle, make the required $10^9$ image classifications, $10^3$ each, in a
few weeks; robotically reducing the target list would lead to a proportional
decrease in the required team size.

Systematic detection of rare astronomical objects by such ``crowd-sourced''
visual inspection has recently been achieved by the online citizen science
project PlanetHunters \citep{SchwambEtal2012}. PlanetHunters was designed to
enable the discovery of transiting exoplanets in data taken by the Kepler
satellite; a community of N inspectors from the general public found, after
each undergoing a small amount of training, N new exoplanet candidates by
visual inspection of the Kepler lightcurves that were presented in a custom
web-based classification interface. The older Galaxy Zoo morphological
classification project \citep{LintottEtal2008} has also enabled the discovery
of rare objects, via its flexible inspection interface and discussion forum
\citep{LintottEtal2009}. Indeed, several of us (AV,EB,CC,TJ,CM,LW) were active
in an informal Galaxy Zoo gravitational lens search, an experience which led
to the present hypothesis that a systematic online visual lens search could be
successful.

In this paper, we describe the \sw website, an online system that enables 
crowd-sourced gravitational lens detection by inviting volunteers to classify
astronomical survey images as containing lens candidates or not.  In a
companion paper we will present the new gravitational lenses discovered in our
first experimental lens search, and begin to investigate the differences
between lens detections made in \sw and those made with automated techniques.
Here though, we try to answer the following questions:

@@

\item How reliably can we find gravitational lenses using the \sw
system? What is the completeness of the sample produced?

\item How noisy is the system? What is the purity of the sample
produced?

\item How quickly can lenses be detected, and non-lenses be rejected?
How many classifications, and so how many volunteers are needed per target?

\item What can we learn about the scalability of the crowd-sourcing approach?

@@

This paper is organised as follows.  In \Sref{sec:design} we introduce the \sw
classification interface and the volunteers who make up the \sw
collaboration,  explain how we use training images, and describe our two stage
candidate selection strategy. We then briefly introduce, in \Sref{sec:data}
the particular dataset used in the first experimental tests of the \sw system,
and how we prepared the images prior to displaying them in the web interface.
In \Sref{sec:swap} we describe our methodology for interpreting the
classifications made by the volunteers, and then present the results of system
performance tests  made on the training images in \Sref{sec:results}. % In
this section, we also investigate the properties of the crowd, to % understand
where the information is coming from. We discuss the implications of our
results for future lens searches in \Sref{sec:discuss} and draw conclusions in
\Sref{sec:conclude}.

% \section{Introduction}
% \label{sec:intro}
% \section{Experiment Design}
% \label{sec:design}
% \subsection{Training}
% \label{sec:design:training}
% \subsection{Stage 1: Initial Classification}
% \label{sec:design:stage1}
% \subsection{Stage 2: Refinement}
% \label{sec:design:stage2}
% \section{Data}
% \label{sec:data}
% \subsection{The CFHT Legacy Survey}
% \label{sec:data:CFHTLS}
% \subsection{Image Presentation}
% \label{sec:data:display}
% \section{Classification Analysis}
% \label{sec:swap}
% \section{Results}
% \label{sec:results}
% \subsection{Crowd Properties}
% \label{sec:results:crowd}
% \subsection{Sample completeness and purity}
% \label{sec:results:sample}
% \section{Discussion}
% \label{sec:discuss}
% \section{Conclusions}
% \label{sec:conclude}

@@

\section{Experiment Design}
\label{sec:design}

The basic steps of a visual search for gravitational lenses are: 1) prepare
images, 2) display them to an inspector, 3) record the inspector's
classification of each image (as, for example, containing a lens candidate or
not) and 4) analyzing those classifications (and all others) in order to
produce a final candidate list. We describe step 1 in \Sref{sec:data} and step
4 in \Sref{sec:swap}. In this section we take a volunteer's eye view and
begin by describing the \sw classification interface, the crowd of volunteers,
and the interactions between the two.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering\includegraphics[width=0.9\linewidth]{sw-system-figs/sw-screengrab-marker+feedback.png}
\caption{Screenshot of the \sw classification interface.}
\label{fig:screenshot}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Classification Interface}
\label{sec:design:interface}

A screenshot of the \sw classification interface (CI) is shown in
\Fref{fig:screenshot}. The CI is the centrepiece of the \sw website,
\texttt{http://spacewarps.org}; the web application is written in
coffeescript, css and html and follows the general design of others written by
the Zooniverse team.\footnote{The \sw web application code is open source and
can be accessed from \texttt{https://github.com/zooniverse/Lens-Zoo}}  The
focus of the CI is a large display of the current pre-prepared PNG image of
the ``subject'' being inspected.   When the image is clicked on by the
volunteer, a marker symbol appears where the pointer was. Several markers can
be placed.  The next image moves rapidly in from a queue formed at the right
hand side of the screen when the ``Finished marking'' button is pressed. At
the same time, the positions of the markers are written out to the
classification database, in an entry that also stores the ID of the subject,
the username (or IP address) of the volunteer, a timestamp and some other
metadata.

Gravitational lenses are rare: typically, most of the images will not contain
a lens candidate, and these need to be quickly rejected by the inspector. The
queue allows several images to be pre-loaded while the volunteer is
classifying the current subject, and the rapid movement is designed to
encourage volunteers to classify rapidly.

For the more interesting subjects, the CI offers two features that  enable
further investigation of the subjects. First is the ``Quick Dashboard'' (QD) a
more advanced image viewer. This allows the viewer to compare three different
contrast and color balance settings, to help bring out subtle features, and to
zoom in on interesting regions of the image to assess small features. Markers
can be placed in the Quick Dashboard just the same as in the main CI image
viewer. The second is a link to that subject's page in the project discussion
forum, \texttt{http://talk.spacewarps.org}. Here, volunteers can discuss the
features they have seen either before they submit their classification, or
after, if they ``favorite'' the subject. There is no ``back'' button: each
volunteer may only classify a given subject once. However, the presence of an
option to see what others think about any given subject before submitting your
own classification means that the classifications may not be  strictly
independent; the advantage of this system is that volunteers can learn from
others what constitutes a good lens candidate. In practice, we might expect
this to be a relatively unimportant educational resource, given the explicit
training we provide for the volunteers, which we describe in the next section.

\subsection{Training}
\label{sec:design:training}

Gravitational lenses are unfamiliar objects to volunteers who are new
to the site. New volunteers need to learn what lenses look like as quickly as
possible, so that they can contribute informative classifications. They also
need to learn what lenses do not look like, in order to reduce the false
positive detection rate. There are three primary mechanisms in the \sw system
for teaching the volunteers what to look for. These are, in the order in which
they are encountered, an inline tutorial, instant feedback,
and a ``Spotter's Guide.''

\subsubsection{Inline Tutorial}

New volunteers are welcomed to the site with a very short tutorial, in which
the task is introduced, a typical image containing a simulated lens is
displayed, and the marking procedure walked through, using pop-up message
boxes. Subsequent images gradually introduce the more advanced
features of the classification interface (the QD and Talk buttons), also using
pop-up messages. The tutorial was purposely kept as short as possible so as to
provide the minimal barrier to entry.

\subsubsection{Instant Feedback}

The second image viewed after the initial tutorial image is already a survey
image, in order to get the volunteers engaged in the real task as quickly as
possible. Training continues beyond the first image tutorial through
``training subjects'' inserted randomly into the stream.  These training
subjects are either simulated lenses (known as ``sims''), or survey images
that were expert-classified (by AV, AM and PM) and found not to contain any
lens candidates (these images are known as ``duds''). The tutorial explains
that the volunteers  will be shown such training images. They are also
informed that they will receive instant feedback about their performance after
classifying (blind) any of these training subjects. Indeed, after a volunteer
finishes marking a training subject and hits ``Finished marking,'' a pop-up
message is generated, containing either positive feedback for a successful
classification (for example, ``Well done! You spotted a simulated lens,'' as
in \Fref{fig:screenshot}) or negative feedback for an unsuccessful one (for
example, ``There is no gravitational lens in this field!'')

\question{PJM}{Does the feedback lead to volunteers leaving the site more 
quickly, or less quickly? Does negative feedback have more impact than
positive?}

The initial frequency of the training images is set to be two in five;
subjects are drawn randomly from the pool of training images with this
frequency. The pool contains equal numbers of sims and duds, and the draw is
made without replacement (for that volunteer). As the number of
classifications made by a volunteer increases, this frequency is decreased, to
$2/(5\times2^{(\textrm{int}(N_c/20)+1)/2})$ ($\approx 0.3$ for the second 20
subjects, $0.2$ for the third 20 subjects, and so on).

This training regime means that in the first 60 images viewed, each volunteer
is shown (on average) 9 simulated gravitational lenses, and 9 empty fields. 
This is a much higher rate than the natural one: to try and avoid this leading
to over-optimism among the inspectors (and a resulting high false positive
rate), we display the current ``Simulation Frequency'' on the classification
interface (``1 in 5'' in \Fref{fig:screenshot}) and maintain the consistent
theme in the feedback messages that lenses are rare.

\subsubsection{Spotter's Guide}

The instant feedback provides real-time educational responses to the
volunteers as they start classifying; as well as this dynamic system, \sw
provides a static reference work for volunteers to consult when in doubt about
how to perform the task. This ``Spotter's Guide'' is a set of webpages showing
example lenses, both real and simulated, and also some common false positives,
drawn from the pool of survey images. The non-lenses were identified by three
of us (AV, AM and PM) while inspecting a small set of survey images in order
to define the ``dud'' training images. For easy reference, the lenses are
divided by type (for example, ``lensed galaxies,'' ``lensed quasars'' and
``cluster lenses''), as are the false positives (for example, ``Rings and
Spirals,'' ``Mergers,'' ``Artifacts'' and so on). The example images are
accompanied by explanatory text. The Spotter's Guide is reached via a button
on the left hand side, or the hyperlinked thumbnail images of the ``Quick
Reference'' provided on the right hand side, of the classification interface.

Most of the text of the Spotter's Guide focuses on what lenses do or don't do;
the website ``Science page'' contains a very brief introduction to how
gravitational lenses work, which is fleshed out a little on the ``FAQ'' page. 
This also contains answers to frequently asked questions about the interface
and the task set.

\subsection{Stage 1: Initial Classification}
\label{sec:design:stage1}

Interface fast due to pre-loading of images, and minimizing interaction.
Trade-off between speed and accuracy. Decreasing training rate.

Quick dashboard provides simple ways to explore further: zoom, contrast
controls.

Spotting lenses: Markers to be placed. Two reasons: first, to give good
feedback. Second, to focus attention.

Non-lenses marked? Favourite button instead, enabling serendipitous
discovery of other interesting things, separate from lenses.

Retirement of low probability systems. Concentrates sample, provides more
``bacon'' (while slightly skewing "sim frequency"). Note that this feature
means that everyone contributes to detection of lenses: luck is made for the
few that happen to see the new lenses, by the masses that did the rejection.
Group effort.

Sims vs duds leads to inclusive search -- click on anything you think etc...

\subsection{Stage 2: Refinement}
\label{sec:design:stage2}

Goal: assess candidates, reject false positives by comparing with training set
of non-obvious non-lenses. Produce a sample rankable by probability.

Reconfigured website: more detailed SG, more detailed feedback. Orange
background to make it obvious stage 2 is different. Slower image presentation.
Higher, constant training rate.

@@

\section{Classification Analysis}
\label{sec:swap}

In this section we outline our methodology for interpreting the interactions
of the volunteers with the identification interface.  Each classification made
is logged in a database, storing subject IDs, (anonymous) volunteer IDs, a
timestamp and the classification results.  The {\it kind} of subject --
whether it is a training subject (a  simulated lens or a known non-lens) or a
test subject (an unseen image drawn from the survey) -- is also recorded. For
all subjects, the positions of all Markers are recorded, in pixel coordinates.
For training subjects, we also store the ``classification'' of the subject as
a lens, or a non-lens, and also the type of object present in the image. These
types are summarized in \Tref{tab:objecttypes}.  This classification is used
to provide instant feedback, but is also the basic measurement used in a
probabilistic classification of every subject based on all image views to
date.

We perform an online analysis of the classifications,  updating a
probabilistic model of every (anonymous) volunteer's data, and also updating
the lens probability of each subject  (in both the training and test sets), on
a daily basis. This gives us a dynamic estimate of the posterior probability
for  any given  subject being a lens, given all classifications of it to date.
Assigning thresholds in this lens probability then allows us to make good
decisions about whether or not to retire a subject from the system, in order to
focus attention on new images.

The details of how the lens probabilities are calculated are given in
\Aref{appendix:swap}. In summary:
\begin{itemize}

\item Each volunteer is assigned a simple software agent, characterised by a
confusion matrix. The two independent elements of this matrix are the
probabilities, as estimated by the agent, that the volunteer is going to be 1)
correct when they report that an image contains a lens when it really does
contain a lens, $\pr(\saidLENS|\LENS,T)$, and 2) correct when they report that
an image does not contain a lens when it really doesn't contain a lens,
$\pr(\saidNOT|\LENS,T)$.

\item Each agent updates its confusion matrix elements based on the number of
times its volunteer has been right in each way while classifying subjects from
the training set, accounting for noise early on due to small number
statistics: $T$ is the set of all training images seen to date.

\item Each agent uses its confusion matrices to update, via Bayes' theorem,
the probability of an image from the test set containing a lens,
$\pr(\LENS|C,T)$, when that image is classified by its volunteer. ($C$ is the
set of all classifications made of this subject.)

In the next section we present results in terms of all three of these
probabilities, as we investigate the performance of the \SW system.

@@

In this section we present our findings about the performance of the  \sw
system, in terms of the information  contributed by the crowd in
\Sref{sec:results:crowd}, and the overall classifications of the training set
that they made.

@@

\subsection{Crowd Properties}
\label{sec:results:crowd}

% Enthusiasm: histogram of classifications, stage 1 vs stage 2. Information
% contributed. Correlation with number of classifications. 
% 
% PL and PD as measures of skill. not quite talent, due to possibility of
% learning - but agents assume talent. Performance of crowd re PD and PL.
% Correlations with N classifications.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering\includegraphics[width=0.9\linewidth]{sw-system-figs/all_skill_contribution_experience_education.png}
\caption{Key properties and contributions of the \sw crowd. Plotted are the
1-D and 2-D marginalized distributions for the logarithms of the 
properties of the agents described in the text. The stage 1 agents are shown
in blue, the stage 2 agents in orange.}
\label{fig:crowd:cornerplot}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We define the following properties of the crowd, as characterised by their
agents, and plot the distributions of their logarithms in
\Fref{fig:crowd:cornerplot}. 
%
\begin{description}
%
\item{\noindent\bf ``Effort:''} The number of test images, $\effort$, classified by a
volunteer. In stage 1, the mean effort per agent was 263; in the shorter stage
2 it was 81.
%
\item{\noindent\bf ``Experience:''} The number of training images, $\experience$,
classified by a volunteer. In stage 1, the mean experience per agent was 29;
in stage 2 (where the training image fequency was set higher) it was 34.
%
\item{\noindent\bf ``Skill:''} The expectation value of the information gain, $\skill$
should  the next subject classified have lens probability 0.5
\Aref{appendix:swap},  in bits. Random classifiers have $\skill = 0.0$,
perfect classifiers have $\skill = 1.0$. All agents start with $\skill = 0.0$;
this increases as training subjects are classified, and the agent's estimates
of its confusion matrix elements improve. In stage 1, the mean skill per agent
was 0.04 bits; in  stage 2 it was 0.05.
%
\item{\noindent\bf ``Contribution:''} The integrated skill over a volunteer's test
subject classification history, and represents the total  contribution to the
project that volunteer (see the appendix for more discussion of this
quantity). In stage 1, the mean contribution per agent was 34.3 bits; in 
stage 2 it was 33.5.
%
\item{\noindent\bf ``Information:''} The total information $\information$ generated by
the  agent during the volunteer's classification activity. This quantity
depends on the value of each subject's lens probability when that subject was 
presented to the volunteer (\Aref{appendix:swap}), and so there is an element
of luck involved with this quantity: if you never see a high probability
subject, it's hard to generate a large amount of information. You make your
own luck by classifying more subjects.
%
\end{description}

The leftmost column of \Fref{fig:crowd:cornerplot} shows how the last four of
these properties depends on the effort expended by the volunteers. We see that
experience is strongly correlated with effort (as training images are
presented throughout each stage, albeit at decreasing frequency), and that
this is also true for skill. In the second row of \Fref{fig:crowd:cornerplot}
we see that while skills of greater than 0.1 can be attained after just a few
training images, most agents of such low experience have significantly lower
skill. The volunteers in question only classify a few subjects before leaving.
However, at high values of experience and effort, the skill is \emph{always
high}. There seem to be very few agents logging large numbers of
classifications at low skill (although there are one or two exceptions). Most
high effort ``super-users'' have high skill. These two properties are
reflected in both the contributions these volunteers make (third row) and the
information they generate (fourth row).

The distributions for the stage 2 agents (orange) are qualitatively similar to
those for the stage 1 agents (blue). Differences are: 1) the maximum effort
possible at stage 2 is smaller, because fewer subjects were available to be
classifed, but 2) the mean effort expended at stage 2 was higher (perhaps
because the subjects were higher probability, and more compelling); 3) the
information generated per agent was higher at stage 2, because the subjects
had higher probability.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=0.9\linewidth]{sw-system-figs/crowd_contrib_cumul.png}
\caption{Narrow cumulative distributions of the contributions made by the
agents: for example, 90\% of the stage 2 contributions were made by the
highest contributing  7\% of the crowd. The stage 1 agents are shown in blue,
the stage 2 agents in orange. ``Experienced volunteers'' classified 10 or more
training subjects.}
\label{fig:crowd:cumulplot}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\Fref{fig:crowd:cornerplot} shows the \sw crowd to have quite broad
distributions of logarithmic effort, skill, and contribution. To better
quantify the contributions made by the volunteers, we show their cumulative
distribution on a linear scale in \Fref{fig:crowd:cumulplot}. This plot shows
clearly the importance of the hardest-working, most active volunteers: at
stage 1,  1.0\% of the volunteers -- 375 people -- made 90\% of the
contribution.  At stage 2, where it was not possible to make as
many classifications before running out of subjects, 7.2\% of the volunteers
-- 141 people -- made 90\% of the contribution.

However, it is not the case that only these small groups were capable of
making this large contribution: 9097 volunteers were ``experienced'' in that
they had all classified at least 10 training images; the 375 highest
contributing volunteers make up just 4\% of this experienced volunteer pool.
The cumulative distribution of agent skill is shown in
\Fref{fig:crowd:cumulskillplot}: these distributions are significantly broader
than the corresponding distributions of agent contribution in 
\Fref{fig:crowd:cumulplot}. The most skilled 20\% of agents possess only  79\%
of the skill at stage 1, and 77\% at stage 2. The inexperienced volunteers
also possess a significant fraction of the skill: the most skillful 20\% of
experienced volunteers (1820 people)  possess just 43\% of the total skill.
The level of contribution made at \sw by experienced volunteers is largely a
matter of choice (or perhaps, availability of time!).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=0.9\linewidth]{sw-system-figs/crowd_skill_cumul.png}
\caption{Broad cumulative distributions of agent skill: 
the most skilled 20\% of the crowd only possess 79\% of the skill at stage 1.
The stage 1 agents
are shown in blue, the stage 2 agents in orange. ``Experienced volunteers''
classified 10 or more training subjects.}
\label{fig:crowd:cumulskillplot}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=0.9\linewidth]{sw-system-figs/stage2_veteran_contribution.png}
\caption{Stage 2 agent properties, separating veterans from stage 1 from new
volunteers. Point size represents total information generated, dashed lines
are drawn at the mean values for each sample.}
\label{fig:crowd:stage1vstage2}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We now look at the transition between stage 1 and stage 2 in more detail. Of
the 1964 volunteers who took part in the stage 2 classification round, only
774 were veterans from stage 1. However, these volunteers showed significantly
higher skill levels than the new users, and also made more classifications
(and hence made greater contributions. This is illustrated in  
\Fref{fig:crowd:stage1vstage2}. In this figure, point size is proportional to
information generated: the higher skill, higher effort stage 1 veterans
generated most of the information.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering\includegraphics[width=0.9\linewidth]{sw-system-figs/early_vs_final_skill.png}
\caption{Predictability of final skill from early skill (after 10 training
classifications) in stage 1 agents. The dashed lines enclose 649 agents with
early skill greater than 0.1 
who retain final skill of greater than 0.05.}
\label{fig:crowd:skillprediction}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Having seen the importance of the high skill volunteers at both stage 1 and
stage 2, we might ask, can we predict agent skill after, say, the volunteer
has experienced 10 training subjects? In \Fref{fig:crowd:skillprediction} we
plot this ``early skill'' against final skill, for all experienced stage 1
volunteers. The majority of the crowd shows little correlation, but at the
high early skill end, some predictability appears. The 649 experienced  stage
1 volunteers who had early skill greater than 0.1 went on to attain a mean
final skill of 0.22, with 97\% remaining at skill 0.05 or higher. This
suggests that it might be worth tracking volunteers' skill as a project
progresses, in order to encourage those showing an aptitude for the task to
take part in the more difficult activities, such as stage 2 refinement.

\Tref{tab:crowd:contributions} shows the total effort, contribution, skill and
information generated in both stage 1 and 2 of the CFHTLS project, with the
total numbers of agents and subjects for comparison. These numbers allow us to
quantify the efficiency of the system. The contribution per classification is
defined in terms of a hypothetical subject with lens probability of 0.5; one
bit of information is needed to update such a subject's lens probability to
either zero or one. This means that a maximally complete classification stage
would yield a total contribution (summed over all agents) equal to the number
of subjects. The ratio of this hypthetical optimum to the actual total
contribution is therefore a measure of the stage's inefficiency. We find our
inefficiency (by dividing column 2 by column 3) to be 33\% and 17\% in stage 1
and 2 respectively. In stage 1, this inefficiency is due to the daily
processing: we were not able to retire subjects fast enough, and so they,
remained in the system, being over-classified. Only 3705704 classifications
were needed to retire all the subjects: the ratio of this to the total number
made is 34\%. (The remaining 1\% is due to not all subjects being classified
to 1 or 0 probability.) At stage 2, we did not retire any subjects at all; the
inefficiency in this case was by design, to give everyone a chance to
appreciate what they had found together!

The total skill of the crowd, computed by summing the skill of all the agents,
is, similarly, a measure of the effective crowd size: a crowd of perfect
classifiers would be this size. The stage 1 crowd was equivalent to a team of
1469.9 perfect classifiers; the stage 2 crowd was equivalent to a team of
102.4 perfect classifiers.

The total information generated during the survey is harder to interpret,
because it's exactly the high probability candidates that we leave in the
system that give rise to most of the information generated. What we can do is
divide the total information generated by the amount of information it takes
to classify a \sw subject all the way to the detection threshold (lens
probability 0.95), divide by this number, and then multiply by the survey
inefficiency to get a very rough estimate for the effective number of
detections corresponding to the crowd's contribution. With a prior probability
of $2\times10^{-4}$, it takes 11.6 bits to make a detection; the effective
number of detections contributed by the stage 1 and stage 2 volunteers is then
2670 and 24 respectively. These figures are close to the numbers of detections
given in column 7 of \Tref{tab:crowd:contributions}. 
The uncertainty in the interpretation of the
information generated provides further justification for our focus on the
expected information gain as a measure of volunteer contribution.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}
\begin{center}
\caption{Total crowd and subject sample properties from the CFHTLS project.}
\label{tab:crowd:contributions}
\begin{tabular}{cccccccc}
  \hline
  \hline {Stage} & Subjects & Contribution                          & Agents & Skill      & Classifications          & Candidates & Information \\
                 & $\Ns$    & $\sum_k^{\Nv} \contribution_k$ (bits) & $\Nv$  & $\sum_k^{\Nv} \skill_k$ (bits) & $\sum_k^{\Nv} \thiseffort$ & $\Ncands$  & $\sum_j^{\Ns}\sum_k^{\Nv} \information_{j,k}$ (bits) \\
  \hline 
            1    & $427064$ & $1291997.4$ & $36929$ & $1469.9$ & $10768759$ & $3367$ & $91110.7$ \\
            2    & $3679$   &   $21895.8$ &  $1964$ &  $102.4$ &   $224745$ &   $90$ &  $1640.4$ \\
  \hline \hline
\end{tabular}
\medskip\\
\end{center}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Sample completeness and purity}
\label{sec:results:sample}

Rejection rate. Completeness and purity at P > retirement, P > 95\%, and 
as function of probability P. Compare stage 1 and stage 2.

Summarize performance at some fiducial threshold: eg P = 95\%.

@@

Challenges for future.

@@

Summary of system.

Crowd-sourced gravitational lens detection works, in terms of the
classification of the training set as described here, in the following
specific ways:

@@

\appendix

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Space Warps CFHTLS Results Paper
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\title[\sw]
{\SW: New Lens Candidates from the CFHT Legacy Survey}

\author[More et al.]{%
  \input{sw-cfhtls-authors.tex}
}

@@

We present X new strong gravitational lens candidates found in the CFHT Legacy
survey by the \sw collaboration.

@@

In this paper, we ... try to answer the following questions:

@@

\item ...

\item ...

\item ...

@@

This paper is organised as follows. In \Sref{sec:sw} we give a brief overview
of the \sw system, focusing on the aspects most relevant to the interpretation
of the results of this first lens search. In \Sref{sec:data} we introduce the
CFHTLS imaging data. ... We discuss the implications of our results for future
lens searches in \Sref{sec:discuss} and draw conclusions in
\Sref{sec:conclude}.

@@

\section{\sw Overview}
\label{sec:sw}

@@

\subsection{Known Lenses}
\label{sec:data:knownlenses}

@@

Completeness and purity with regard to known lenses.

@@

\subsection{Stage 1 Classification}
\label{sec:results:stage1}

@@

\subsection{Stage 2 Classification}
\label{sec:results:stage2}

\subsection{Comparison with ``Expert'' Classification}
\label{sec:results:experts}

\subsection{New Lens Candidates}
\label{sec:results:newlenses}

@@

Differences with robots: what types of lenses are found at \sw?

Selection function. Missing system.

Further work.

@@

Summary of CFHTLS project.

Crowd-sourced gravitational lens detection works, in terms of the
classification of the CFHTLS image dataset as described here, in the following
specific ways:

@@

% \appendix
